---
title: "Meshing Uluru"
author: "Miles McBain"
output: html_document
---

```{r}
library(sf)
library(raster)
library(tidyverse)

source("./tris_sf/helpers/sf_to_tri_mesh.R")
```
# Meshing Uluru
The akward scale of the North Carolina dataset is probably atypical of what we will be exporting to VR. Here I'll look at meshing Uluru, which should be more suitable.

Data sources: 
[Uluru / Ayers Rock 1:250 000 GIS Dataset](http://pid.geoscience.gov.au/dataset/89644)
[ELVIS - Elevation information system](http://www.ga.gov.au/elvis/)

```{r}
uluru_bbox <- st_bbox(c(ymin = -25.36109, ymax = -25.32889, xmax = 131.0747, xmin = 131.0117), crs = st_crs("+proj=longlat +ellps=GRS80 +no_defs"))
uluru_shape <- read_sf("./tris_sf/data/GEODATA_TOPO250K_TILE_DATA/G5208/Shapefiles/Elevation/g5208_contours.shx")
uluru_contours <- 
    filter(uluru_shape, as.vector(st_contains(st_as_sfc(uluru_bbox), uluru_shape$geometry, sparse = FALSE)))

# Those interpolated features are kind of weird.
plot(uluru_contours["FEATTYPE"])

# What if we just got the outline

# Find the outer contour by finding the one with the highest longitude?
uluru_outline <- 
  uluru_contours %>%
  mutate(max_y = map_dbl(geometry, ~max(as.matrix(.)[,1]))) %>% # [,1] gives longitude
  arrange(desc(max_y)) %>% 
  filter(row_number() == 1)

uluru_contours_bbox <- st_bbox(uluru_contours)

uluru_outline_poly <- 
  uluru_outline %>%
  st_geometry() %>%
  .[[1]] %>%
  as.matrix() %>%
  list() %>%
  list() %>%
  st_multipolygon() %>%
  st_sfc()

st_crs(uluru_outline_poly) <- st_crs(uluru_contours)
uluru_outline_poly <- st_transform(uluru_outline_poly, 
'+proj=lcc +lat_1=-30 +lat_2=-20 +lat_0=-25 +lon_0=135 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0') #same as raster below

uluru_trimesh <- sf_to_tri_mesh(uluru_outline_poly, 5000)
plot(uluru_trimesh)
```

# Adding Elevation
Okay so that has worked pretty well. now we try to add some elevation stuff.

We have two approaches: 
1. Build an elevation raster from the elevation contours.
2. Use ELVIS 1 Second Shuttle Elevation data

## Use ELVIS 1 Second elevation raster
This has the slightly different crs. So I made it the same via `st_transform` in the previous step.

```{r}
nt_raster <- raster("./tris_sf/data/ELVIS_CLIP.tif")

# sanity check - do these overlap?
plot(nt_raster)
plot(as(uluru_outline_poly, 'Spatial'), add = TRUE)
# looks okay but doe it fit?

uluru_elev <- raster::extract(nt_raster, uluru_trimesh$P[,1:2])
uluru_trimesh$P <- cbind(uluru_trimesh$P, uluru_elev)

library(rgl)
rgl.clear()
  wire3d(
    tmesh3d(vertices = t(asHomogeneous(uluru_trimesh$P)), indices = array(t(uluru_trimesh$T)))
  )
rglwidget()
```
The front side looks a bit whacky. Comparse this to elevation of a square section.

```{r}
bbox_to_multipoly <- function(a_bbox, crs){
  points_mat <- rbind(
      c(a_bbox$xmin, a_bbox$ymin),
      c(a_bbox$xmax, a_bbox$ymin),
      c(a_bbox$xmax, a_bbox$ymax),
      c(a_bbox$xmin, a_bbox$ymax),
      c(a_bbox$xmin, a_bbox$ymin))
  colnames(points_mat) <- NULL
  
  mpoly <- st_sfc(st_multipolygon(
  list(list(
      points_mat  
    ))
  ))
  st_crs(mpoly) <- crs
  names(mpoly) <- NULL
  mpoly
}

uluru_extent <- 
  bbox_to_multipoly(uluru_contours_bbox, "+proj=longlat +ellps=GRS80 +no_defs")

uluru_extent <-
  st_transform(uluru_extent, 
    ' +proj=lcc +lat_1=-30 +lat_2=-20 +lat_0=-25 +lon_0=135 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0') # same as raster

uluru_box_trimesh <- sf_to_tri_mesh(uluru_extent, 5500)
plot(uluru_box_trimesh)

uluru_box_elev <- raster::extract(nt_raster, uluru_box_trimesh$P[,1:2])
uluru_box_trimesh$P <- cbind(uluru_box_trimesh$P, uluru_box_elev)

rgl.clear()
  wire3d(
    tmesh3d(vertices = t(asHomogeneous(uluru_box_trimesh$P)), indices = array(t(uluru_box_trimesh$T)))
  )
rglwidget()
```
